---
title: "Analysing effect of abiotic factors on intercept of mixed effect model"
author: "Petra Guy"
date: "11 May 2018"
output: pdf_document
---
Her the intercept from the mixed model is anaysed. The data is split in the same way, and outliers from area and PHI are removed.

```{r setup, include=FALSE}

## NB - aseveral libraries mask each other here - arm masks dplyr and corrplot, therefore open libraries #as required.
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
# Richness modelling

rm(list = ls())
cat("\014")
library(dplyr) # everything

library(ggplot2)
library(gridExtra)
library(car) # for vif
library(reshape) # melt

# get the data
site_data =  read.csv("../../Data/CompleteSiteLevelVars.csv")
nestZs = readRDS("../nest_mixed_model_fits.RDS")
ints = nestZs%>%select(Site,intercept)
zeta_r = readRDS("../Zeta/zeta_r")
Site = c(1:103)
zeta_r = as.data.frame(cbind(Site,zeta_r))
site_data = inner_join(site_data,zeta_r)

#mean impute the missing PHI
meanPHI = round(mean(site_data$Pos_Hetero_Index, na.rm = TRUE),2)
x = site_data$Pos_Hetero_Index
x[is.na(x)] = meanPHI
site_data$Pos_Hetero_Index = x

site_data_ints = inner_join(site_data,ints)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
correl = round(cor(nestZs$intercept, nestZs$slope, method = "spearman"),1)
g1 = ggplot(nestZs, aes(x = intercept, y = slope))+
  geom_point()+
  geom_smooth(method = lm)+
  labs(title =paste("spearman ", correl))+
  ylab("slope of log log fit")+
xlab("intercept of log log fit")

g2 = ggplot(site_data_ints, aes(x = intercept, y = Richness ))+
  geom_point()+
  geom_smooth(method = lm)+
  xlab("intercept of log fit")
 



data = nestZs%>%select(Site,slope,intercept)
data$cz = data$slope*data$intercept
richness = site_data_ints%>%select(Site, Richness)

slopeint = inner_join(richness,data)

barfill <- "#4271AE"
barlines <- "#1F3552"

g3 = ggplot(slopeint, aes(x = slope, y = Richness ))+
  geom_point()+
  geom_smooth(method = lm)+
  xlab("z")

grid.arrange(g1,g2,g3, ncol=3)
```

Having a quick look at the behaviour of the intercepts and slopes, as richness increases, slope and intercept increase, but the first graph shows that as the slope increases the intercept decreases. I have no idea what this means, if anything.

```{r}
barfill <- "#4271AE"
barlines <- "#1F3552"
 hist = ggplot(slopeint, aes(x = slopeint$cz))+
  geom_histogram(binwidth = 0.05, colour = barlines, fill = barfill)+
   xlab(NULL)+
 theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

plot = ggplot(slopeint, aes(x = cz, y = Richness))+
        geom_point()+
        geom_smooth( method = lm)+
        ylab("Richness")+
        ggtitle("Richness with cz")


#(hist,plot, nrow = 2, heights=c(2,5))
plot
```

Since the gradient is a function of c and z, I thought it might be nice to look at richness with cz. But I have no idea what this means either.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# slect only required variables
subset_all = site_data_ints%>%select("Site","intercept","Area_ha",
                                "Northing", "Pos_Hetero_Index","Buffer3",
                                "no_MSG", "no_NVC","sd_pH","sd_SOM","sd_LBA",
                                "sd_meandbh","sd_treedensity","area_ratio",
                                "meandbh","meanph", "meanSOM","meanLBA",
                                "meantreedensity","zeta_r")



colnames(subset_all) = c("Site","ints","Area",
                         "Northing", "PHI","Buffer",
                         "no_MSG", "no_NVC","sd_pH","sd_SOM","sd_LBA",
                         "sd_meandbh","sd_TD","area_ratio",
                         "meandbh","meanph", "meanSOM","meanLBA",
                         "meanTD","zeta_r")
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#remove the wood with the largest area
largest_area = as.numeric(subset_all%>%filter(Area == max(Area))%>%select(Site))
site_data_outlier1 = subset_all%>%filter(Site!=largest_area)
site_data_outlier1 = site_data_outlier1[,-3] # remove area column now

#remove the outlier in PHI
largest_PHI = as.numeric(subset_all%>%filter(PHI == max(PHI))%>%select(Site))
site_data_outlier2 = site_data_outlier1%>%filter(Site!=largest_PHI)

largest_PHI = as.numeric(site_data_outlier2%>%filter(PHI == max(PHI))%>%select(Site))
site_data_outlier3 = site_data_outlier2%>%filter(Site!=largest_PHI)

```



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
subset_sd = site_data_outlier3%>%select("Site","ints",
                               "Northing", "PHI","Buffer",
                               "no_MSG", "no_NVC","sd_pH","sd_SOM","sd_LBA",
                               "sd_meandbh","sd_TD","area_ratio","zeta_r")



subset_mean = site_data_outlier3%>%select("Site","ints",
                              "Northing", "PHI",  "meandbh",
                              "meanph", "Buffer", "meanSOM","meanLBA",
                              "meanTD","area_ratio", "no_NVC", 
                              "no_MSG","zeta_r")

```




```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width = '50%'}

# we know richness vs ph usually unimodal around .5, therefore fit to meanpH^2
data = subset_mean[,-1]
SDs = subset_mean[,2]
data$meanph = (data$meanph)^2

#rescale the data
library(arm) #for standarize
rescaled_mean_data = apply(data[,-1],2, rescale)
rescaled_mean_data = as.data.frame(cbind(SDs, rescaled_mean_data))
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#create the model
mod_mean = lm(SDs~., data=rescaled_mean_data, na.action = "na.fail")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#have a look at the linear model
par(mfrow =c(2,2))
plot(mod_mean, main = "Mean dataset")
```
The two site with the highest values of PHI had high leverage in this model and were therefore removed from the data in order to give normally distributed residuals, the plots above were created after these values were removed.
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width = '50%'}

# we know richness vs ph usually unimodal around .5, therefore fit to meanpH^2
data = subset_sd[,-1]
ints = subset_mean[,2]

#rescale the data
library(arm) #for standarize
rescaled_sd_data = apply(data[,-1],2, rescale)
rescaled_sd_data = as.data.frame(cbind(ints, rescaled_sd_data))
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#create the model
mod_sd = lm(ints~., data=rescaled_sd_data, na.action = "na.fail")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#have a look at the linear model
par(mfrow =c(2,2))
plot(mod_sd, main = "SD dataset")
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# look at vifs
vif(mod_mean)
```
The variance inflation factors in the mean dataset are low, suggesting that correlations between covariates are low and not likely to increase the variance of the paramater estimates.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# look at vifs
vif(mod_sd)
```

The variance inflation factors in the sd dataset are also low


The first twelve models from the mean dataset, which had a delta <2 were selected from the MuMin dredge funtion as the top model set. 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(MuMIn) #dredge and avg
 #get top models
models = dredge(mod_mean)
model_set = get.models(models, subset = delta<2)


#do model averaging, subset means zero method
mean_avg_models = model.avg(model_set, subset)

#select output data
summary = summary(mean_avg_models)
coefs =  mean_avg_models$coefficients
importance =  c(NA,(as.vector(mean_avg_models$importance)))

coef_matrix = summary$coefmat.full
coefs = coef_matrix[,1]
adj_se = coef_matrix[,3]
CI_lower =  coefs - 1.96*adj_se
CI_upper = coefs + 1.96*adj_se

output = round(as.data.frame(cbind(coefs,adj_se, CI_lower, CI_upper, importance)),2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# make plot of the variables and CI
n = nrow(output)
data = output[c(2:n),]
data = data[,c(1,3,4)]
data = t(data)
data = as.data.frame(data)
melted = melt(data)

fun_mean <- function(x){
  data = data.frame(y=mean(x),label=mean(x,na.rm=T))
  data = round(data,2)
  return(data)}

#(aes_string(x = 'variable', y='value', na.rm = TRUE)

melted$variable = as.factor(melted$variable)# ps, wont plot as separate plots if x continous

ggplot(melted,aes_string(x = 'variable', y ='value' ) )+
  geom_boxplot(na.rm = TRUE, width = 0)+
  stat_summary(fun.y = mean, geom="point",colour="darkred", size=3) +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7)+
  stat_summary(geom = "text", label = output$importance[-1],fun.y = max, hjust = 1, colour = "red")+
  geom_abline(intercept = 0, slope = 0)+
  labs(y = "Effect size and 95%CI",x = "effect")+
  labs(title = "Model averaged results for delta <2, intercepts, Mean dataset",
       subtitle = "numbers in red are variable importance")
  


```

The graph shows the averaged effect sizes of the model with delta < 2.The buffer and meanpH may effect the intercep of the log/log fit

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(MuMIn) #dredge and avg
 #get top models
models = dredge(mod_sd)
model_set = get.models(models, subset = delta<1.5)


#do model averaging, subset means zero method
sd_avg_models = model.avg(model_set, subset)

#select output data
summary = summary(sd_avg_models)
coefs =  sd_avg_models$coefficients
importance =  c(NA,(as.vector(sd_avg_models$importance)))

coef_matrix = summary$coefmat.full
coefs = coef_matrix[,1]
adj_se = coef_matrix[,3]
CI_lower =  coefs - 1.96*adj_se
CI_upper = coefs + 1.96*adj_se

output = round(as.data.frame(cbind(coefs,adj_se, CI_lower, CI_upper, importance)),2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# make plot of the variables and CI
n = nrow(output)
data = output[c(2:n),]
data = data[,c(1,3,4)]
data = t(data)
data = as.data.frame(data)
melted = melt(data)

fun_mean <- function(x){
  data = data.frame(y=mean(x),label=mean(x,na.rm=T))
  data = round(data,2)
  return(data)}

#(aes_string(x = 'variable', y='value', na.rm = TRUE)

melted$variable = as.factor(melted$variable)# ps, wont plot as separate plots if x continous

ggplot(melted,aes_string(x = 'variable', y ='value' ) )+
  geom_boxplot(na.rm = TRUE, width = 0)+
  stat_summary(fun.y = mean, geom="point",colour="darkred", size=3) +
  stat_summary(fun.data = fun_mean, geom="text", vjust=-0.7)+
  stat_summary(geom = "text", label = output$importance[-1],fun.y = max, hjust = 1, colour = "red")+
  geom_abline(intercept = 0, slope = 0)+
  labs(y = "Effect size and 95%CI",x = "effect")+
  labs(title = "Model averaged results for delta <1.5,interepts, SD dataset",
       subtitle = "numbers in red are variable importance")
  


```
The Northing is the only variable which may effect the intercepts of the log/log fit


##Using the model for prediction
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#extract model averaged parameter estimates

predicted_int = predict(mean_avg_models, full = TRUE)
empirical_int =  site_data_outlier3$ints
fit = lm(empirical_int ~ predicted_int)
R2 = round(summary(fit)$r.squared,2)
f = summary(fit)$fstatistic
p = round(pf(f[1],f[2],f[3], lower.tail = F),4)
subtitle = paste("R2 = ",R2, "p = ",p)
data = as.data.frame(cbind(predicted_int, empirical_int))

ggplot(data, aes(x = predicted_int, y = empirical_int))+
  geom_point()+
  geom_abline(intercept = 0, slope = 1)+
  labs(title = "Observed versus predicted data, mean dataset",
       subtitle = subtitle)

  
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#extract model averaged parameter estimates

predicted_int = predict(sd_avg_models, full = TRUE)
empirical_int =  site_data_outlier3$ints
fit = lm(empirical_int ~ predicted_int)
f = summary(fit)$fstatistic
p = round(pf(f[1],f[2],f[3], lower.tail = F),4)
subtitle = paste("R2 = ",R2, "p = ",p)
data = as.data.frame(cbind(predicted_int, empirical_int))

ggplot(data, aes(x = predicted_int, y = empirical_int))+
  geom_point()+
  geom_abline(intercept = 0, slope = 1)+
  labs(title = "Observed versus predicted data, sd dataset",
       subtitle = subtitle)

  
```



```{r}
h = slopeint$Richness
y = slopeint$slope
x = slopeint$intercept

scatter3D(x, y, h,  colvar = h, col = NULL, add = FALSE,
xlab = "intercept",
ylab = "slope",
zlab = "Richness",
phi = 20, theta = 20)
  

```