---
title: "ExaminingCovariatesII"
author: "PetraGuy"
date: "27 March 2018"
header-includes: 
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output: pdf_document
---
##Repeat of covariance analysis without sites 23, 53 , 74

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	out.width="50%"
)
rm(list = ls())
cat("\014")
library(dplyr)
library(visreg) # for visreg plots 
library(ggplot2)
library(GGally) # for ggpairs
library(factoextra)
library(FactoMineR)# these two for PCA
library(corrplot)
```


```{r}
sitevars =  read.csv("../Data/SiteLevelVars.csv")
toremove = c(23,53,74)
sitevars = sitevars%>%filter(!Site %in% toremove)
```

```{r}
physical_df = sitevars%>%select("Alt_m", "Area_ha","Perim_m","Easting","Northing","Buffer3","meandbh","meanph","meanSOM","meanLBA","area_ratio")
hetero_df = sitevars%>%select("Pos_Hetero_Index","no_NVC","sd_pH", "sd_SOM" , "no_MSG", "sd_LBA", "sd_meandbh", "sd_treedensity","no_trees"  )
hetero_df[is.na(hetero_df)] = 0
```

Physical Variables
```{r}
colnames(physical_df)
```

Heterogeneity variables

```{r}
colnames(hetero_df)
```

#Effect of each variable on richness using multiple linear regression, simple additive model.


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
physical_df_incRich = as.data.frame(cbind(sitevars$Richness, physical_df))
colnames(physical_df_incRich) = c("Richness","Alt_m" ,"Area_ha" ,"Perim_m" ,"Easting","Northing" ,"Buffer3" , "meandbh" , "meanph" ,"meanSOM" , "meanLBA","area_ratio" )
physical_model = lm(Richness ~., data = physical_df_incRich)

visreg(physical_model)
```



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
heter_df_incRich = as.data.frame(cbind(sitevars$Richness, hetero_df))
colnames(heter_df_incRich) = c("Richness","Pos_Hetero_Index","no_NVC", "sd_pH" ,"sd_SOM", "no_MSG", "sd_LBA" ,"sd_meandbh" , "sd_treedensity" , "no_trees"  ) 
hetero_model = lm(Richness ~., data = heter_df_incRich)

visreg(hetero_model)
```


\clearpage
\newpage
\blandscape

#Colinearity of variables using pair plots

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="250%"}

ggpairs(physical_df, axisLabels="none",
        lower = list(continuous="smooth"),
        diag="blankDiag",
        upper = list(corSize=2,axisLabels='show'))+
theme(legend.position = "none",
      panel.grid.major = element_blank(),
      axis.ticks = element_blank(),
      panel.border = element_rect(linetype = "dashed", colour = "black", fill = NA))

```



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE,out.width="250%"}

ggpairs(hetero_df, axisLabels="none",
        lower = list(continuous="smooth"),
        diag="blankDiag",
        upper = list(corSize=2,axisLabels='show'))+
theme(legend.position = "none",
      panel.grid.major = element_blank(),
      axis.ticks = element_blank(),
      panel.border = element_rect(linetype = "dashed", colour = "black", fill = NA))
```
\clearpage
\newpage
\elandscape


#PCA
##Physical variables

```{r}
physical_PCA = PCA(physical_df, scale.unit = TRUE, ncp = 11, graph = FALSE)
var = get_pca_var(physical_PCA)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
eig.val_physical = get_eigenvalue(physical_PCA)
eig.val_physical
```



```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
fviz_eig(physical_PCA, addlabels = TRUE, ylim = c(0,30))
```
    
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_pca_var(physical_PCA, col.var = "cos2", repel = TRUE)
```

 
```{r echo=FALSE, message=FALSE, warning=FALSE}

corrplot(var$cos2)
```



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_cos2(physical_PCA, choice = "var")
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
fviz_contrib(physical_PCA, choice = "var", axes = 1, top = 10)
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
fviz_contrib(physical_PCA, choice = "var", axes = 2, top = 10)

```


```{r eval=FALSE, include=FALSE}
area_outier = sitevars%>%filter(Area_ha == max(Area_ha))


```

```{r}
hetero_PCA = PCA(hetero_df, scale.unit = TRUE, ncp = 11, graph = FALSE)
var = get_pca_var(hetero_PCA)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
eig.val_hetero = get_eigenvalue(hetero_PCA)
eig.val_hetero
```



```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
fviz_eig(hetero_PCA, addlabels = TRUE, ylim = c(0,30))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_pca_var(hetero_PCA, col.var = "cos2", repel = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
corrplot(var$cos2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
fviz_cos2(hetero_PCA, choice = "var")
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
fviz_contrib(hetero_PCA, choice = "var", axes = 1, top = 10)
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
fviz_contrib(hetero_PCA, choice = "var", axes = 2, top = 10)

```

#Discussion

Removing the outliers in area and positive heterogeneity index has:
Reduced slope of richness ~ pos hetero index
Redued the contribution of altidude to component 4, (Site 74 is Glen Beasdale, could be up on the hill).
Area contribution now more distributed towards component 2.



