---
title: "ExaminingCovariatesI"
author: "PetraGuy"
date: "27 March 2018"
header-includes: 
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	out.width = "50%"
)
rm(list = ls())
cat("\014")
library(dplyr)
library(visreg) # for visreg plots 
library(ggplot2)
library(GGally) # for ggpairs
library(factoextra)
library(FactoMineR)# these two for PCA
library(corrplot)
```


```{r}
sitevars =  read.csv("../Data/SiteLevelVars.csv")
```
Several approaches will be used to examine the collineartiy of the variates used in the to explain he richness of the 16 plots in each site, and the response of the richness to the variates will be considered,

A) Visref will be used to plot scatter graphs of individual variables used in a simple additive multiple linear regression.
B) Pairs plots will be used to examine collinearity between all variables.
C) PCA will be used to see if we can extract a few key variables to represent the data.
D) VIR values will be calculated
E) p values for the variables will be recorded as variables are removed from the regression.
F)ANOVA will be used to see if there is a difference between the means

Since there are a lot of variables, they have been split in two, those which directly relate to abiotic features and those which represent the heterogeneity of the site:

```{r}
physical_df = sitevars%>%select("Alt_m", "Area_ha","Perim_m","Easting","Northing","Buffer3","meandbh","meanph","meanSOM","meanLBA","area_ratio")
hetero_df = sitevars%>%select("Pos_Hetero_Index","no_NVC","sd_pH", "sd_SOM" , "no_MSG", "sd_LBA", "sd_meandbh", "sd_treedensity","no_trees"  )
hetero_df[is.na(hetero_df)] = 0
```

Physical Variables
```{r}
colnames(physical_df)
```

Heterogeneity variables

```{r}
colnames(hetero_df)
```

#Effect of each variable on richness using multiple linear regression, simple additive model.

Using visref to look at the variables allows a multiple regression to be carried out and the effect of each variable is shown with the other variables held constant. This allows for the effect of each variable on the richness, in the presence of the other variables to be considered.

```{r}
physical_df_incRich = as.data.frame(cbind(sitevars$Richness, physical_df))
colnames(physical_df_incRich) = c("Richness","Alt_m" ,"Area_ha" ,"Perim_m" ,"Easting","Northing" ,"Buffer3" , "meandbh" , "meanph" ,"meanSOM" , "meanLBA","area_ratio" )
physical_model = lm(Richness ~., data = physical_df_incRich)

visreg(physical_model)
```


The plots show that the following variables appear to influence the species richness:
Northing, Buffer, mean dbh, mean SOM, area to perimeter ratio.


```{r}
heter_df_incRich = as.data.frame(cbind(sitevars$Richness, hetero_df))
colnames(heter_df_incRich) = c("Richness","Pos_Hetero_Index","no_NVC", "sd_pH" ,"sd_SOM", "no_MSG", "sd_LBA" ,"sd_meandbh" , "sd_treedensity" , "no_trees"  ) 
hetero_model = lm(Richness ~., data = heter_df_incRich)

visreg(hetero_model)
```

The plots show that the following variables appear to influence the species richness:
Number of positive heterogeneity indices (Although this may be heavily influenced by the two outliers, Site 23 with 97 and Site 53 with 78), number of NVC codes, number of sites with no trees, standard devitation of the tree density, (weak).

\clearpage
\newpage
\blandscape

#Colinearity of variables using pair plots

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="250%", fig.cap= "The pairs plots show no correlation greather than 0.52 between any variables, accept area and perimeter and area ratio - as you'd expect. Buffer and Northing have correlation coefficient of 0.52 and Easting and mean SOM of  -0.417."}

ggpairs(physical_df, axisLabels="none",
        lower = list(continuous="smooth"),
        diag="blankDiag",
        upper = list(corSize=2,axisLabels='show'))+
theme(legend.position = "none",
      panel.grid.major = element_blank(),
      axis.ticks = element_blank(),
      panel.border = element_rect(linetype = "dashed", colour = "black", fill = NA))

```


\clearpage
\newpage

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="250%", fig.cap="The largest correlation coefficient is 0.385, between standard deviation of mean dbh and number of sites with no trees, not unexpected. Number of major soil groups and standard devation of SOM have correlation coefficient of 0.337 and number of NVC codes and number of plots with no trees has 0.299."}

ggpairs(hetero_df, axisLabels="none",
        lower = list(continuous="smooth"),
        diag="blankDiag",
        upper = list(corSize=2,axisLabels='show'))+
theme(legend.position = "none",
      panel.grid.major = element_blank(),
      axis.ticks = element_blank(),
      panel.border = element_rect(linetype = "dashed", colour = "black", fill = NA))
```

\elandscape
\clearpage
\newpage

#PCA
##Physical variables

```{r}
physical_PCA = PCA(physical_df, scale.unit = TRUE, ncp = 11, graph = FALSE)
var = get_pca_var(physical_PCA)
```

```{r}
eig.val_physical = get_eigenvalue(physical_PCA)
eig.val_physical
```

5 components explain 76% of the variance, the first 2 explain 45%

```{r}
fviz_eig(physical_PCA, addlabels = TRUE, ylim = c(0,30))
```
    
```{r}
fviz_pca_var(physical_PCA, col.var = "cos2", repel = TRUE)
```

 
```{r}

corrplot(var$cos2)
```



```{r}
fviz_cos2(physical_PCA, choice = "var")
```

```{r}
fviz_contrib(physical_PCA, choice = "var", axes = 1, top = 10)
```

```{r}
fviz_contrib(physical_PCA, choice = "var", axes = 2, top = 10)

```

Area, perimeter, buffer contribute to the first component, area ratio and areA to the second, mean dbh and mean LBA to the third, northing and altitUde to the fourth.

The PCA suggests including these variables would explain 68% of the variance.

The multiple linear regression shows a positive correlation of Northing, buffer, and mean SOM to richness, and a neative correlation with mean dbh. Area did not have a strong correlation with richness.  

The strong contribution of area and perimeter may be due to the outlier - one wood had a area of over 300ha,(Site 74) the others were all below 150ha. Site 74 will be removed and the analysis repeated.


The inclusion of area, perimeter and area/periemter ratio ay seems strange, but they may have different effects on richness. A large area may increase richnes due to increased species pool. A large area/perimeter ratio may increase richness due to the increased possibility of species migration, or conversely have a negative impact because the wood may then not contain a core of undisturbed woodland with poorly dispersing woodland species, whcih are then lost. Buffer may be a good substitute for all the above variables because it also represents a species pool, a wood with a large perimeter and a large area will equally have a large buffer. 

```{r}
area_outier = sitevars%>%filter(Area_ha == max(Area_ha))


```

```{r}
hetero_PCA = PCA(hetero_df, scale.unit = TRUE, ncp = 11, graph = FALSE)
var = get_pca_var(hetero_PCA)
```

```{r}
eig.val_hetero = get_eigenvalue(hetero_PCA)
eig.val_hetero
```

75% of variance is expressed by the first 5 components,  41% by the first 2

```{r}
fviz_eig(hetero_PCA, addlabels = TRUE, ylim = c(0,30))
```

```{r}
fviz_pca_var(hetero_PCA, col.var = "cos2", repel = TRUE)
```

```{r}
corrplot(var$cos2)
```

```{r}
fviz_cos2(hetero_PCA, choice = "var")
```

```{r}
fviz_contrib(hetero_PCA, choice = "var", axes = 1, top = 10)
```

```{r}
fviz_contrib(hetero_PCA, choice = "var", axes = 2, top = 10)

```

The standard deviation of soil organic matter, tree density, LBA and mean dbh contribute the first principle component. Number of NVC codes, sd of soil pH and number of sites with no trees to the second and number of MSG to the third. The corr plot shows sd SOM increases with sd of tree density and sd of mean dbh, and that number of NVC codes increase with sd of soil pH.

The PCA suggests including sd SOM, sd mean dbh, sd tree density, number NVC, sd pH, and number NVC codes would explain 54% of the variance, adding positive heterogeneity index would increase that to around 75%
The multiple inear regression shows that Positive hetero index, number of NVC and number of plots with no trees correlate positively with richness.

The correlation of positive hetero indices with richness may be due to the two outliers, Site 53 and site 23. Analysis will be repearted without these.

