---
title: "Exploring Nests"
author: "PetraGuy"
date: "1 March 2018"
output: pdf_document
---
```{r setup, include=FALSE, message=FALSE, comment=FALSE,error=FALSE}
# MUCH OF THIS IS A REPEAT OF NestLevelData coz I wanted things for my talk.
# Have added map of sites and bit to look at pread of richness in woods and calcuation of plot richness
# INPUTS ground cover and vegetation BRC codes
# OUTPUTS plot_rich.csv df of all sites/plots - richness of each full plot in each site
knitr::opts_chunk$set(echo = FALSE)
rm(list = ls())
cat("\014")
library(dplyr)
library(ggplot2)
library(ggmap)
library(ggrepel)
library(nlme)
library(reshape)
library(gridExtra)

```

These analysis will take the vegetation data (Groundcover.csv) and compute species richness by nest,plot and site. These values will be saved to csv files which can be added to the Nest and Plot level data sets.
```{r}
#the wood location data


woods = read.csv("../Data/EastingNorthing.csv")
colnames(woods) = c("Site", "Easting","Northing","GridRef","Lat","Long")

#just for plotting
data = cbind(woods$Lat, woods$Long)
```

```{r, comment=FALSE,message=FALSE}
#Input all the ground flora data
Data = read.csv("../Data/GroundCover.csv")
Data_Yr2 = Data%>%filter(Yr_2 == 2)#%>%select(SITE,PLOT,NEST,COV,Amalgams)
colnames(Data_Yr2) = c("SITE", "PLOT","NEST","Cover","BRC_number","Year")
veg_codes = read.csv("../Data/vegetation_codes.csv")

# the  bryophytes, lichen etc have already been removed from these, so the veg codes csv files is no bryophytes, so need to join this with ground cover to eliminate bryophytes from counts


colnames(veg_codes) = c("Species", "BRC_number")
Data_Yr2_veg = Data_Yr2%>% inner_join(veg_codes)
# now using Dat_Yr2_veg means analysis is carried out without bryophytes
```


```{r}
# returns the species richness for each site/plot/nest 
# NAs are not counted, these occur for sapling counts which need to be added in to the plot richness.

nest_richness = function(data){
  site_list = list()
  #browser()
    for (i in 1:103){
    sitedata = data%>%filter(SITE == i)
    plot_nest_df = data.frame()
    for (j in 1:16){
      plotdata = sitedata%>%filter(PLOT == j)
      tmp = vector()
       for (k in 1:5){
        nestdata = plotdata%>%filter(NEST ==k)
        tmp = c(tmp, nrow(nestdata))
        }
      plot_nest_df = rbind(plot_nest_df, tmp)
      colnames(plot_nest_df) = c("nest1","nest2","nest3","nest4","nest5")
    }
   site_list[[i]]= plot_nest_df
  }
return(site_list)   
}
spec_rich = nest_richness(Data_Yr2_veg)
saveRDS(spec_rich,"SpecRichness.RDS")
```
Since nest level data is only modelled as species~area - no other vars, spec_rich list is all thats needed. This data will be analysed from this document, a csv is not needed because it is not being added #to another file.spec_rich is list, **spec_rich[[1]]** returns **list** of richnesses for plot 1-16, nest 1-5 in site 1 this does not include saplings unless recored - i.e. they will be little baby plants.

```{r}
# now need to include the NAs and combine them into a plot richness
plot_richness = function(data){
  site_matrix = matrix(nrow = 103, ncol = 16)
  #browser()
  for (i in 1:103){
    sitedata = data%>%filter(SITE == i)
    plot_nest_df = data.frame()
    for (j in 1:16){
      plotdata = sitedata%>%filter(PLOT == j)
      site_matrix[i,j] = length(unique(plotdata$BRC_number))


    }
  }
  return(site_matrix)
}
plot_rich = plot_richness(Data_Yr2_veg)

write.csv(plot_rich,"../Data/plot_rich.csv")
```

For the seedlings there is sometimes record in specified plot, and NA record, therefore have  used unique to capture those that are NAs, but not double count when same species listed in a nest. **plot_rich[i,j]** returns **dataframe** of richness for site i plot j. May not equate to sum of nest richness because saplings are included.

Missing plots come up as 0s in order not to be counted in frequencies- but note, some "missing" plots do have vegetation entries, and these are counted. Handle within each woodland modelled, and replace with NAs when fitting models.

Species richness per nest is not saved to csv because its a list.
use spec_rich[[1]] for nest richness of all plots of site 1

```{r}
# just calculates richness found in each wood, no extrapolation.
#d[i] is richness of wood i
basic_richness = function(data){
  d = rep(0,103)
 for (i in 1:103){
   site = data%>%filter(SITE==i)
   d[i]  = length(unique(site$BRC_number))
 }
  return(d)
} 
d = basic_richness(Data_Yr2_veg)
#write.csv(d,"../Data/SiteRichness.csv") - this done in NestLevelData

```

**d[1]** will give **vector** of total sum of plot richnesses across the site, with repeated BRCs codes removed.

```{r}
## cumulative richness for site i, for all 16 plots.
##need spec_rich calculated above

cum_rich_all = list()
for ( i in 1:103){
  #browser()
  site = spec_rich[[i]]
  site = as.data.frame(site)
  cum_rich_site = data.frame()
   for (j in 1:16) {
    plot = site[j,]
    r = 0
    cumrich = rep(0,5)
    for (n in 1:5){
      r = r + plot[[n]]
      cumrich[n] = r
    }
    cum_rich_site = rbind(cum_rich_site, cumrich)
   }
  colnames(cum_rich_site) = c("nest1","nest2","nest3","nest4","nest5")
  cum_rich_all[[i]] = cum_rich_site
  }
saveRDS(cum_rich_all, "CumulateveRichness.RDS")
```

**cum_rich_all** is a **list** of cumulative richness for each plot/site. cum_rich_all[[1]] gives cumulative richness for all plots in site 1 for nest 1 - 5



```{r}
#some stats, might use these
wood_rich = cbind(woods,d) # just add richness to locations

max_wood = wood_rich%>%filter(Site==match(max(d),d))

min_wood = wood_rich%>%filter(Site==match(min(d),d))
#wont work with 0s
plot_rich[plot_rich == 0] = NA
min_max = rbind(min_wood, max_wood)
mins = apply(plot_rich,1,min,na.rm=TRUE)
maxs = apply(plot_rich,1,max,na.rm=TRUE)
ranges = maxs - mins
max_range= max(ranges)
min_range = min(ranges)
max_range_wood = wood_rich%>%filter(Site==match(max(ranges),ranges))
min_range_wood = wood_rich%>%filter(Site==match(min(ranges),ranges))
stats = rbind(max_wood,min_wood, max_range_wood, min_range_wood)
values = c(max(d),min(d),max_range,min_range)
n = colnames(stats)
n = c(n[-7],"values")
colnames(stats) = n
#rownames(stats) = c("max richness","min richness","max range","min range")

```

**stats** is dataframe of wood locations containing maximum and minimum richness and richness ranges. Other stats of interest cann be added to this to create summary data.


```{r}
#Look at top and bottom quartiles?
u_woods = wood_rich%>%filter(d>100)
l_woods = wood_rich%>%filter(d<60)
good_bad = rbind(u_woods,l_woods) #good_bad used in mappin below

SOM = site_data$meanSOM
wood_rich$SOM = SOM
```

```{r, message=FALSE, warning=FALSE}
#Some maps like look nice
#uses good_bad and data
#ggmap version,  scaled these by diversity with max richess/min richness ####

map<- get_map(location = c(lon = -1.5, lat = 54),
                         color = "color",
                         source = "google",
                         maptype = "satellite",
                         zoom = 6) 
#png("../Data/Talk/Map.png") 
ggmap(map, extent ='device')+
    geom_point(aes(x = Long, y = Lat, colour = SOM),data = wood_rich, size = 2, alpha=1)+
    scale_color_gradient(low = "yellow", high="red")+
  geom_point(aes(x = Long, y = Lat, colour = d),data = min_max, size = 4, alpha=1)+
  scale_color_gradient(low = "yellow", high="red")+
  geom_text_repel(data = good_bad, aes(x = Long, y = Lat,label = d), 
            size = 3, colour = "white")

##################################

ggmap(map, extent ='device')+
  geom_point(aes(x = Lon, y = Lat, colour = meanSOM),data = woods_SOM, size = 4, alpha=1)+
  scale_color_gradient(low = "yellow", high="red")

```




```{r, warning = FALSE, message = FALSE}
#select nearby woods
#uses data

nearby_woods = woods%>%filter(Long> -1.3 & Lat < 51.5  )

map<- get_map(location = c(lon = 0, lat = 51.25),
              color = "color",
              source = "google",
              maptype = "satellite",
              zoom = 9) 

ggmap(map, extent ='device')+
  geom_point(aes(x = Long, y = Lat),data = nearby_woods, color="red", size=2, alpha=1)+
   geom_text_repel(data = nearby_woods, aes(x = Long, y = Lat,label = GridRef), 
            size = 3, colour = "white")

###############################

```

```{r, messgase=FALSE, error=FALSE, warning=FALSE, comment=FALSE}
#function to create a list of graphs of raw cf from the vector sitevector

plots = c("plot1","plot2","plot3","plot4","plot5","plot6","plot7",
          "plot8","plot9","plot10","plot11","plot12","plot13","plot14","plot15","plot16")
area = c("4","25","50","100","200") 
areas = sort(as.numeric(rep(area,16)), decreasing = FALSE)


plot_site = function(sitevector){
 #browser()
  for (i in 1:length(sitevector)){
    site = sitevector[i]
    cf_site = cum_rich_all[[site]]
    cf_site$plot = plots
    melted_cf = melt(cf_site)
    melted_cf$area = areas
    siteno  = paste("Site",site)
    p = ggplot(melted_cf)+
      geom_point(aes(x = area, y = value), colour = "black")+
      labs(title = siteno)+
      ylab("cumulative richness")+
      theme(plot.title = element_text(size = 10))
     plot_list[[i]] = p
     
      
  }
  return(plot_list)
}

# repeat above but plot on A^0.25 axis to show relationship more easily

plot_site_trans = function(sitevector){
 #browser()
  for (i in 1:length(sitevector)){
    site = sitevector[i]
    cf_site = cum_rich_all[[site]]
    cf_site$plot = plots
    melted_cf = melt(cf_site)
    melted_cf$area = areas
    siteno  = paste("Site",site)
    p = ggplot(melted_cf)+
      geom_point(aes(x = (area)^0.25, y = value), colour = "black")+
      labs(title = siteno)+
      ylab("cumulative richness")+
      theme(plot.title = element_text(size = 12))
     plot_list[[i]] = p
     
      
  }
  return(plot_list)
}
```


```{r error=FALSE, message=FALSE, warning=FALSE, comment=FALSE,paged.print=FALSE}
# scatter plots of raw cf data for top 8 woods
up_woods =  wood_rich%>%filter(d>115)
site_to_plot = up_woods$Site
plot_list =list()

plots_all = plot_site_trans(site_to_plot)

#png("../Data/Talk/8Bestcf.png")
grid.arrange(plots_all[[1]], plots_all[[2]], plots_all[[3]],
             plots_all[[4]], plots_all[[5]], plots_all[[6]],
             plots_all[[7]],plots_all[[8]], ncol = 2)
```

```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
#scatter plots of raw cf for worst 8 woods
lo_woods = wood_rich%>%filter(d<45)
site_to_plot = lo_woods$Site
plot_list =list()
plots_all = plot_site_trans(site_to_plot)

#ng("../Data/Talk/8worst.png")
grid.arrange(plots_all[[1]], plots_all[[2]], plots_all[[3]],
             plots_all[[4]], plots_all[[5]], plots_all[[6]],
             plots_all[[7]],plots_all[[8]], ncol = 2)
```

There follows some exploration of the best wood, site 83, richness 150.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# modelling the best wood only
best_wood = wood_rich[wood_rich$d == max(wood_rich$d),]
x = best_wood$Site

f_bestsite = spec_rich[[x]]
cf_bestsite = cum_rich_all[[x]]
f_bestsite[f_bestsite==0]=NA
cf_bestsite[cf_bestsite==0]=NA
#f_bestsite = f_bestsite[!is.na]
melted_f = melt(f_bestsite)
melted_cf = melt(cf_bestsite)


# particular prep adding plots, removing plot 16 - for plot 83 empty - and remelting
plots = c("plot1","plot2","plot3","plot4","plot5","plot6","plot7",
          "plot8","plot9","plot10","plot11","plot12","plot13","plot14","plot15")
cf_bestsite = cf_bestsite[-16,]
cf_bestsite$plot = plots
melted_cf = melt(cf_bestsite)
f_bestsite = f_bestsite[-16,]
melted_f = melt(f_bestsite)

#lets add areas to data.

area = c("4","25","50","100","200") 
areas = sort(as.numeric(rep(area,15)), decreasing = FALSE)
melted_cf$area = areas
melted_f$area = areas

```

```{r}
#box plots of best wood, data from above chunk
melted_f$area = as.factor(melted_f$area) # ps, wont plot as separate plots if x continous
ggplot(melted_f, (aes_string(x='area', y='value', na.rm = TRUE)) )+
   geom_boxplot(na.rm = TRUE, outlier.shape = 7, outlier.size = 3)+
   ggtitle("Site 83, best wood, spread of richness per nest across all plots")

#box plots now just space out neatly, not to scale, so cant put on same graph
```

```{r}
# nlme log log model for best wood NB log = ln

model_nlme_area_log = lme(log(value)~log(area),random = ~1|plot, data = melted_cf, na.action = na.omit)
fit3 = predict(model_nlme_area_log)
melted_cf$fit3 = fit3
```

```{r, error=FALSE, message=FALSE, warning=FALSE}
#model plot for best wood
ggplot(melted_cf)+
  geom_point(aes(x=log(area),y=log(value),na.rm =TRUE), colour = "black") +
  geom_line(aes(x=log(area), y = fit3, group = plot), colour = "red")+
  geom_abline(intercept = 2.2497,slope = 0.2357, size = 2, colour = "blue")+
  ggtitle("Site 83")+
  ylim(low = -0.6, high = 5)+
         annotate("text",x = 2, y = 4, label  = "S = 9.9A^0.24")
```

```{r, warning=FALSE, error=FALSE, message=FALSE}
#raw data cf scatter plot best wood
ggplot(melted_cf)+
 geom_point(aes(x = area, y = value), colour = "black")+
  ggtitle("Raw data, site 83, cumulative richness across all plots with increasing area")+
  annotate("text", x = 75, y = 60, 
           label = "Change in richness due to area less than change in richness due to plot")

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#look at species in site 83
site83 = Data_Yr2_veg%>%filter(SITE==83)
site83distinctBRC = distinct(site83,BRC_number)
species_list = inner_join(site83distinctBRC, veg_codes)
#png("../Data/Talk/richlist")
grid.table(species_list)
species_list
```



```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# modelling the worst wood only
worst_wood = wood_rich[wood_rich$d == min(wood_rich$d),]
x = worst_wood$Site

f_worstsite = spec_rich[[x]]
cf_worstsite = cum_rich_all[[x]]
f_worstsite[f_worstsite==0]=NA
cf_worstsite[cf_worstsite==0]=NA
#f_bestsite = f_bestsite[!is.na]
melted_f = melt(f_worstsite)
melted_cf = melt(cf_worstsite)


# particular prep adding plots, removing plot 16 - for plot 83 empty - and remelting
plots = c("plot2","plot4","plot5","plot6","plot7",
          "plot8","plot9","plot10","plot11","plot12","plot13","plot14","plot15","plot16")
cf_worstsite = cf_worstsite[-c(1,3),]
cf_worstsite$plot = plots
melted_cf = melt(cf_worstsite)
f_worstsite = f_worstsite[-c(1,3),]
melted_f = melt(f_worstsite)

#lets add areas to data.

area = c("4","25","50","100","200") 
areas = sort(as.numeric(rep(area,14)), decreasing = FALSE)
melted_cf$area = areas
melted_f$area = areas

```

```{r}
#box plots of worst wood, data from above chunk
melted_f$area = as.factor(melted_f$area) # ps, wont plot as separate plots if x continous
ggplot(melted_f, (aes_string(x='area', y='value', na.rm = TRUE)) )+
   geom_boxplot(na.rm = TRUE, outlier.shape = 7, outlier.size = 3)+
   ggtitle("Site 78, worst wood, spread of richness per nest across all plots")

#box plots now just space out neatly, not to scale, so cant put on same graph
```



```{r, warning=FALSE, error=FALSE, message=FALSE}
#raw data cf scatter plot worst wood
#png("../Data/Talk/8Worst.png")
ggplot(melted_cf)+
 geom_point(aes(x = area, y = value), colour = "black")+
  ggtitle("Raw data, site 78, cumulative richness across all plots with increasing area")+
  annotate("text", x = 100, y = 9, 
           label = "Change in richness due to area less than change in richness due to plot")+
  ylim(0,10)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#look at species in site 83
site78 = Data_Yr2_veg%>%filter(SITE==78)
site78distinctBRC = distinct(site78,BRC_number)
species_list = inner_join(site78distinctBRC, veg_codes)
#png("../Data/Talk/poorlist")
grid.table(species_list)
species_list
```


This modelling reverts to all woods - NB, same df names used, so bestwood data now overwritten. **fits** is a **dataframe** of intercepts and coefficients from nlme ln/ln model with nests as fixed and plot as random.

```{r, warning=FALSE, error=FALSE, message=FALSE}
##fit every wood and look at coefficients
#prepare data
plots = c("plot1","plot2","plot3","plot4","plot5","plot6","plot7",
          "plot8","plot9","plot10","plot11","plot12","plot13","plot14","plot15","plot16")
area = c("4","25","50","100","200") 
areas = sort(as.numeric(rep(area,16)), decreasing = FALSE)

fits_all_woods = function(){
  coef_df = data.frame(matrix(ncol = 2, nrow = 0))
  
  #prepare data
  for (i in 1:103){
    #browser()
    cf_site = cum_rich_all[[i]]
    cf_site$plot = plots
    melted_cf = melt(cf_site)
    melted_cf$area = areas
    # zeros will freak this model out, make them NAs
    melted_cf[melted_cf==0]=NA
    #model
    model_nlme_area_log = lme(log(value)~log(area),random = ~1|plot, data = melted_cf, na.action = na.omit)
    
    #get coefficients
    int = model_nlme_area_log$coefficients$fixed[[1]]
    slope = model_nlme_area_log$coefficients$fixed[[2]]
    p_slope = summary(model_nlme_area_log)$tTable[2,5]
    row = c(int,slope, p_slope)
    coef_df = rbind(coef_df,row)
  }
  colnames(coef_df) = c("intercept","slope","p_slope")
  return(coef_df)
}
fits = fits_all_woods()

#max and min slope
maxvars = fits[which(fits$slope == max(fits$slope)),]
minvars = fits[which(fits$slope == min(fits$slope)),]

ggplot()+
  geom_abline(data = fits,aes(intercept = intercept, slope = slope), colour = "pink")+
  geom_abline(data = maxvars,aes(intercept = intercept, slope = slope), colour = "red" )+
  geom_abline(data = minvars,aes(intercept = intercept, slope = slope), colour = "blue" )+
  
  scale_x_continuous(limits = c(0,5))+
  scale_y_continuous(limits = c(0,4))+
  ggtitle("ln/ln fits for all woodlands")+
  annotate("text", x = 3, y = 0.5, 
           label = "maximum slope 0.43, minumum slope = 0.098")+
  labs(x = "ln(area)", y = "ln(richness)")


```

Exploring the best and worst woods.

```{r}
#pdf("../Results/slopes_hist")
ggplot(fits, aes(slope))+
  geom_histogram(binwidth = 0.05, col = "black", fill = "grey")+
  ggtitle("Distribution of slopes in linear mixed effect ln/ln fit", subtitle = "S = cA^0.25 ish")


```

```{r}
data = melted_cf
fit = model_nlme_area_log
#model plot for best wood
ggplot(data, aes(x = log(area), y = log(value), colour = plot )) +
  geom_point() +
  geom_line(aes(y = predict(fit), group = plot))

  
```


```{r}
ggplot(fits, aes(slope))+
  geom_histogram(binwidth = 0.05, col = "black", fill = "grey")+
  ggtitle("Distribution of slopes in linear mixed effect ln/ln fit", subtitle = "S = cA^0.25 ish")


```



```{r, warning=FALSE, error=FALSE, message=FALSE}
##fit every wood and look at coefficients ---adding in the r.squared for each site
#prepare data
plots = c("plot1","plot2","plot3","plot4","plot5","plot6","plot7",
          "plot8","plot9","plot10","plot11","plot12","plot13","plot14","plot15","plot16")
area = c("4","25","50","100","200") 
areas = sort(as.numeric(rep(area,16)), decreasing = FALSE)

fits_all_woods = function(){
  r2_list = list()
  
  #prepare data
  for (i in 1:103){
    #browser()
    cf_site = cum_rich_all[[i]]
    cf_site$plot = plots
    melted_cf = melt(cf_site)
    melted_cf$area = areas
    # zeros will freak this model out, make them NAs
    melted_cf[melted_cf==0]=NA
    #model
    model_nlme_area_log = lme(log(value)~log(area),random = ~1|plot, data = melted_cf, na.action = na.omit)
    
    #get coefficients r.squared
    r2_list[[i]] = r.squaredGLMM(model_nlme_area_log)
  }
  
  return(r2_list)
}
fits = fits_all_woods()
```

```{r}
r_fixed = vector()
r_random = vector()
for (i in 1:103){
  r_fixed[i] = fits[[i]][[1]]
  r_random[i] = fits[[i]][[2]] - fits[[i]][[1]]
  
}
data = as.data.frame(cbind(r_fixed,r_random))

ggplot(data=data, aes(x = r_random, y =  r_fixed))+
  geom_point()+
  geom_abline(slope = -1, intercept = 0)
```
