---
title: "NestSD"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
cat("\014")
library(dplyr)
library(ggplot2)
library(gridExtra)
library(MuMIn)
library(reshape)
library(energy)#dcor
library(corrplot)

library(corrr)

AllSiteVars = read.csv("../Data/CompleteSiteLevelVars.csv")
nestdata = readRDS("nest_mixed_model_fits.RDS")
AllPlotVars= read.csv("../Data/AllPlotsVarsRichness.csv")

```

Repeat of last, now looking at standard deviation of random intercepts obtained from the linear mixed effects model.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
todelete = c(1,3,4,5,6,7,8,9,11,12,14,18, 22,23,24,25,26,27,28,29,30)
Sitevars = AllSiteVars[-todelete]
colnames(Sitevars) = c("Site","PHI","Buffer","Num_NVC","sd_pH", "sd_SOM" ,"sd_LBA",
                        "sd_meanDBH","sd_TD","meanDBH" , "meanph" , "meanSOM" , "meanLBA" ,
                        "meanTD", "area_ratio" )
# have removed the vars I know arent much cop, this time I am going to use the standard deviation variables.
sd_vars = c(1,2,3,4,5,6,7,8,9,15)
Set_sd = Sitevars[sd_vars]

#take site and slope from nestdata
NestSD = nestdata[-c(2,3,4)]
#join to SiteData
SitedataSD = inner_join(Set_sd,NestSD)
SitedataSD[is.na(SitedataSD)] = 0

```



```{r sdhistogram, echo=FALSE, message=FALSE, warning=FALSE, out.width="50%", paged.print=FALSE}

ggplot(NestSD, aes(x = sd_intercept))+
  geom_histogram(binwidth = 0.05, colour = "black",fill = "grey")+
  ggtitle("Distribution of standard deviations of random intercepts")
```
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="75%"}
# first look for response of z to the variables

melted = melt(SitedataSD[-1], id.vars = "sd_intercept")


ggplot(melted, aes(x = value, y = sd_intercept))+
  geom_point()+
  geom_smooth(method = loess)+
    facet_wrap(~variable, scales = "free")+
  scale_x_log10()

```

The scatter plots for each variable show little correlation with sd of intercepts.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
corrs = sort(round(apply(SitedataSD[-c(1,11)],2, function(x) dcor(x,SitedataSD$sd_intercept)), digits = 2), decreasing = TRUE)
corrs = as.data.frame(corrs)
corrs
```

The correlations (using dcor) are low

Dont feel there's much mileage in this...Look at correlation of richness to standard deviation on intercepts..

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
SD_Rich = as.data.frame(cbind(AllSiteVars$Richness, SitedataSD$sd_intercept))
colnames(SD_Rich) = c("Richness","sd_intercept")
ggplot(SD_Rich, aes(x = sd_intercept, y = Richness))+
  geom_point()+
  geom_smooth(method = loess)+
 ggtitle("Pearson and Kendal correlations below 0.2")

corK = cor(AllSiteVars$Richness,SitedataSD$sd_intercept, method = "kendall")
corP = cor(AllSiteVars$Richness,SitedataSD$sd_intercept, method = "pearson")

```

Hmm...moving on, not going to polish a dead horse. Quick look at z's again, this time from averse cumulative SAC across 16 plots.





