---
title: "More Zeta Modelling"
author: "Petra Guy"
date: "26 May 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
rm(list = ls())
cat("\014")

emp_zetas = readRDS("empirical_zetas.RDS")
modelled_coefs = readRDS("zeta_coefficients.RDS")
modelled_zetas = readRDS("modelled_zetas.RDS")

ground_flora = read.csv("../../Data/GroundCover.csv")
ground_flora = ground_flora%>%filter(Yr_2 == 2)
veg_codes = read.csv("../../Data/vegetation_codes.csv")
colnames(ground_flora) = c("SITE", "PLOT","NEST","Cover","BRC_number","Year")
colnames(veg_codes) = c("Species", "BRC_number")
flora = ground_flora%>% inner_join(veg_codes)
Richness = read.csv("../../data/SiteRichness.csv")
Richness = Richness[,-1]

CompleteSiteLevelvars = read.csv("../../Data/CompleteSiteLevelVars.csv")
```

#A coefficient that might be useful??

I suggest: zeta1 and zeta2 are the average alpha and beta diversity of a site - the difference between them tells you something about the heterogeneity of the woodland. 

The gradient of the zeta decline curve indicates the heterogeneity of a site. For a perfectly homogenous site, where all species occur in all plots, all the zeta values are the same and the gradient of the zeta decline curve is zero. For a perfectly heterogenous woodland, each plot has a different set of species, therefore every zeta from order 2 onwards is zero and the gradient of the zeta decline curve is zeta1/1 = zeta1. 

In reality, the site will have an initial gradient somewhere between, but zeta1 - zeta2 could be used as a measure of the heterogeneity of the wood. Although only using zeta1 and zeta2 omits higher orders of zeta, and therefore reduces our knowledge of the heterogeneity on a larger scale, it does include all plots within a site, and therefore considers the site as a whole. This is because the zeta diversity calculation uses every plot to calculate zeta2.  The problem is that this value is not comparable between woodlands because richer woodlands will have a higher zeta1. Therefore consider the ratio r = (zeta1-zeta2)/zeta1. This varies between 1, for a perfectly heterogeneous woodland and 0 for a perfectly homogeneous woodland - regardless of the richness.

Then, if richness is increased with heterogeneity, you will see a positive correlation between site richness's and the coefficient r.

```{r echo=FALSE, out.width="50%"}


r = vector()
for( i in 1:103){
  z = emp_zetas[,i]
  r[i] = (z[1]-z[2])/z[1]
}
c = cor.test(r, Richness, method = "pearson")
pearson = round(c$estimate,2)
p = round(c$p.value, 4)

data7 = as.data.frame(cbind(Richness,r))

ggplot(data = data7, aes(x = r, y = Richness))+
  geom_point()+
  geom_smooth(method = lm)+
  labs(title = "Richness with r for the 103 sites of the woodland survey",
        subtitle = paste("pearson correlation coefficient = ",pearson,"p = ", p))


```

#Notes on zeta modelling and additional

When we tried to fit the exponential or power law to the empirical zeta decay we found a poor fit. Most of the error was in the fitted value for zeta1. The power law tended to overestimate this while the exponential fit underestimated it.

```{r, out.width="50%"}

diff_exp = vector()
diff_pwr = vector()

for (i in 1:103){
  mod_z_exp = modelled_zetas[[1]][1,i]
  mod_z_pwr = modelled_zetas[[3]][1,i]
  z_emp = emp_zetas[1,i]
  diff_exp[i] = z_emp - mod_z_exp
  diff_pwr[i] = z_emp - mod_z_pwr
}
xticklabs = c("exponential model","power law model")

data = as.data.frame(cbind(diff_exp, diff_pwr))
ggplot(stack(data), aes(x = ind, y = values))+
  geom_boxplot(width = 0.3, fill = "grey")+
  ylab("Difference in empirical and modelled values")+
  xlab("Model type")+
  ggtitle("The empircal - modelled value for zeta 1" )+
  scale_x_discrete(labels=xticklabs)


```

This meant that when checking the accuracy of the model by using it to predict the species richness of our 16 plots - the value was either much too low or much too high.

CP suggested a fit of exp(x^-b). This has the problem that it does not decay to zero - as our zeta values often do. But, exp(x^-b)*exp(-ax) does decay to zero due to the second exponential term. The first exponential term is just e for zeta = 1, and therefore with a coefficient multiplier, this expression could be used to increase the estimated value of zeta1 whilst ensuring the values can decay to zero.
So I tried another model:

zeta =  c*exp(x^-a)exp(-bx).

The model was fitted using the nls function.

```{r}

sites_to_remove = c(5,8,24,44,62,64,67,94)
subset_emp_zeta = emp_zetas[,-sites_to_remove]

m = list()
cor = vector()
pred = data.frame()
sites = c(1:4,6,7,9:23,25:43,45:61,63,65,66,68:93,95:103)
mod_coefs = list()
for (i in sites){
  #browser
  colname = paste("Site",i)
  y = subset_emp_zeta[,colname]
  nas = sum(is.na(y))
  l = 16 - nas
  y = y[c(1:l)]
  x = seq(from = 1, to = l, by = 1)
  m[[i]] = nls(y~c*(exp(x^(a)))*exp(b*x), start = list(a = -2, b = -0.1, c = 10))
  p = predict(m[[i]],x)
  cor[i] = cor(p,y, method = "kendall")
  pad = rep(NA,nas)
  p = c(p,pad)
  pred = as.data.frame(rbind(pred,p))
  mod_coefs[[i]] = coef(summary(m[[i]]))
  
  #data1 = as.data.frame(cbind(x,y))
  #data2 = as.data.frame(cbind(x,p))
  #ggplot(data = data1, aes(x = x, y = y) )+
  #geom_point(size = 4)+
  #geom_point(data = data2, aes(x = x,y = p), colour = "red")
}
new_mod_zetas = t(pred)
rownames(new_mod_zetas) = c(1:16)
colnames(new_mod_zetas) = sites
saveRDS(new_mod_zetas,"new_model_zetas")
saveRDS(mod_coefs,"new_mod_coefs")
#this gives pred - the predicted zeta values it is 103x16 instead of 16x103!
```

```{r, out.width="50%"}
pred_zeta1 = pred[,1]
emp_zeta1 = subset_emp_zeta[1,]
diff_newmod = as.data.frame(t(pred_zeta1-emp_zeta1))

# now do boxplot again of diffs

xticklabs = c("New model","Exponential model","Power law model")
data = as.data.frame(cbind(diff_exp, diff_pwr))
ggplot(stack(data), aes(x = ind, y = values))+
  geom_boxplot(width = 0.3, fill = "grey")+
  geom_boxplot(data = diff_newmod, aes(x = "", y = diff_newmod),width = 0.3, fill = "black")+
  ylab("Difference in empirical and modelled values")+
  xlab("Model type")+
  ggtitle("The empircal - modelled value for zeta 1" )+
  scale_x_discrete(labels=xticklabs)

```

The new model is correctly estimating zeta1.

```{r, out.width="50%", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
exp_R2 = vector()
pwr_R2 = vector()
for (i in 1:103) {
  exp_R2[i] = modelled_coefs[[1]][5,i]
  pwr_R2[i] = modelled_coefs[[3]][5,i]
}
  
# get R2 for newmod
#get actual y
subset_emp_zeta = emp_zetas[,sites]

predt = t(pred)

newmod_r2 = vector()
for(i in 1:95){
  y = na.omit(subset_emp_zeta[,i])
  x = predt[,i]
  l = length(y)
  x = x[1:l]
  newmod_r2[i] = summary(lm(y~x))$r.squared
}

###boxplots of R2

data1 = as.data.frame(cbind(exp_R2,pwr_R2))

data3 = as.data.frame(newmod_r2)

ggplot(data = stack(data1), aes(x = ind, y = values))+
  geom_boxplot(width = 0.3, fill = "grey")+
  geom_boxplot(data = data3, aes(x = "",y = newmod_r2),width = 0.3, fill = "grey")+
    scale_x_discrete(labels=xticklabs)+
  ggtitle("R2 values for the three models")+
  ylab("R2")+
  xlab("Model")

```

The R2 suggests the model is good at estimating the values of zeta.

What do the coefficients look like?

```{r, out.width ="50%" ,echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

newmod_coefs = lapply(m, coef)
a = vector()
b = vector()
c = vector()

for (i in 1:103){
  if (!is.element(i,sites_to_remove)) {
  a[i] = newmod_coefs[[i]][[1]]
  b[i] = newmod_coefs[[i]][[2]]
  c[i] = newmod_coefs[[i]][[3]]
  }
  else{
    a[i] = NA
    b[i] = NA
    c[i] = NA
  }
}

saveRDS(a,"a.RDS")
saveRDS(b,"b.RDS")
saveRDS(c,"c.RDS")
#This gives vectors a,b,c for coefficients a,b,c of the model - 
#remember some woods are missing.
## Now look at range in coefficients
xticklabs = c("a","b","c")
data = as.data.frame(cbind(na.omit(a),na.omit(b),na.omit(c)))
ggplot(data = stack(data), aes(x = ind, y = values), na.rm = TRUE)+
  geom_boxplot(width = 0.3, fill = "grey")+
  scale_x_discrete(labels=xticklabs)+
  ggtitle("Values of the coefficients in the model")
```

a and b are very similar across all the models. the upper and lower quartiles for a are -1.6<a<1.01 and for b, -0.4<b<-0.14. (PS - I include the negative here but the model has a negative value for a and b, I dont mean that you substitute -1.6 into exp(x^-b))

The upper and lower quartiles of c are such that 7.28<c<11.16. This large value for c is probably what allows the fit to correct estimate zeta1. For zeta1, exp(x^-a) is 1, exp(-bx) varies between 0.6 and 0.86, therefore by fitting an appropriate value for c, the model can easily result in accurate estimates.

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width="50%", paged.print=FALSE}
c_noNA = na.omit(c) 
zeta1 = t(subset_emp_zeta[1,])
data = as.data.frame(cbind(zeta1,c_noNA))
colnames(data) = c("zeta1","c")

ggplot(data=data, aes(x = c, y = zeta1))+
  geom_point()+
  ggtitle("Zeta1 with c")
```  
The graph shows that c increases with zeta1, such that an accurate fit will be achieved.

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width="50%", paged.print=FALSE}
data = as.data.frame(cbind(na.omit(b), na.omit(a))) 
colnames(data) = c("b","a") 
ggplot(data=data, aes(x = a, y = b))+
  geom_point()+
  ggtitle("b and a")
```

There doesn't appear to be any relationship between b and a.
Are a or b related to richness??

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width="50%", paged.print=FALSE}
library(gridExtra)
Rich = CompleteSiteLevelvars%>%select(Site,Richness)
Rich_subset = Rich%>%filter(!(Site %in% sites_to_remove))
Rich_ordered = Rich_subset[order(Rich_subset$Site),]

data1 = as.data.frame(cbind(Rich_ordered,na.omit(a)))
colnames(data1) = c("Richness","a")
g1 = ggplot(data1, aes(x = a, y = Richness))+
  geom_point()

data2 = as.data.frame(cbind(Rich_ordered,na.omit(b)))
colnames(data2) = c("Richness","b")
g2 = ggplot(data2, aes(x = b, y = Richness))+
  geom_point()

grid.arrange(g1,g2)

```

Not really.

So it seems that the model zeta = cexp(x^-a)exp(-bx) fits the data well, with the values of a and b being very similar across all sites. (PS - 8 sites were omitted because the nls algorithm did not converge using the same starting values that were used on all the other sites). But - what does this new model mean? The power law and the exponential law had the advantage of interpretable ecological meaning; exponential model implying that the probability of occupancy across plots is equal (more homogeneous site??), whereas the power model implies a niche differentiation process  (more heterogeneous site??).

But what does this new model represent? And does it only fit because it has 3 parameters instead of 2?

