---
title: "SiteLevelData"
author: "PetraGuy"
date: "15 March 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#INPUTS from access, ground cover, eg codes, species lists, SiteVars
#       from other progs, posheterogindices, allplotvarsrichness, siterichness, landcoverindices
#OUTPUTS, none - this just for graphing
```

The richness of all 16 plots in each site was explored with various environmental variables. Since all the areas are then equal, the effect on richness is only due to environmental variables.

The first four graphs plot richness against various physical factors:

The site level variables are altitude (Alt_m), Area (Area_ha), perimeter (perim_m), the ratio of area to perimeter (area_ratio), the location (Easting and Northing) and the proportion of positive buffer zones at 1500m, 2500m and 3500m radius around the wood (Buffer1,Buffer2, Buffer3). Positive buffer zones are woodland or natural landscapes, not arable or urban. 
Some plot level variables have been averaged across the plots and are also used, these are mean dbh (meandbh), mean pH (meanph), mean soil organic matter (meanSOM) and mean live basal area (mean LBA).

The next two graphs plot richness again factors which can be taken to represent the heterogeneity of the woodlands:
The number of different NVC codes (no_NVC), the number of positive side codes recorded (Pos_Hetero_Index), the standard deviation of the soil pH and soil organic matter (sd_pH, sd_SOM), the number of different major soil groups recorded (no_MSG), the standard deviation of the live basal area (sd_LBA), standard deviation of the mean dbh  of a plot (sd_meandbh), standard deviation of the tree density of the plots (sd_treedensity) and the number of plots with no trees recorder (no_trees).

The final two plots just use the Ellenberg values. The first plot uses the standard deviation of the Ellenbergs across the plots, the final plot uses the mean values of the Ellenberg for all the plots.

NB - for all the plots I have normalised the x axis to 1, just to make everything fit neatly on the same axis.
I didnt do that for the Ellenberg plots because they are in roughly the same scale anyway


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
rm(list = ls())
cat("\014")
setwd("~/Documents/CMEECourseWork/CMEEMainProject/Code")

Data = read.csv("../Data/GroundCover.csv")
Data_Yr2 = Data%>%filter(Yr_2 == 2)#%>%select(SITE,PLOT,NEST,COV,Amalgams)
colnames(Data_Yr2) = c("SITE", "PLOT","NEST","Cover","BRC_number","Year")
veg_codes = read.csv("../Data/vegetation_codes.csv")

# the  bryophytes, lichen etc have already been removed from these, 
#because the veg codes csv files is no bryophytes, so need to join this 
#with ground cover to eliminate bryophytes from counts


colnames(veg_codes) = c("Species", "BRC_number")
Data_Yr2_veg = Data_Yr2%>% inner_join(veg_codes)

#get the ellenbergs file

ellenbergs = read.csv("../Data/Spp_lib.csv")
ellenbergs  = ellenbergs[c(2,3,5,6,7,8)]
colnames(ellenbergs) = c("BRC_Name","BRC_number","R","N","W","L")
veg_ellens = left_join(Data_Yr2_veg,ellenbergs)
#Now have veg_ellens = site/plot/nest species list and ellenbergs - use this extract
#range of ellenbergs for a Site


#The smple  site richness
SiteRichness = read.csv("../Data/SiteRichness.csv")
colnames(SiteRichness) = c("Site","Richness")
#plot level env vars
AllPlotsvars = read.csv("../Data/AllPlotsVarsRichness.csv")

#positive site desc codes
SiteCodes = read.csv("../Data/PositiveHeterogIndices.csv")

#all other site level vars
SiteLevelVars = read.csv("../Data/SiteVars.csv")

#land cover in buffers
Buffers = read.csv("../Data/LandCoverIndices.csv")
colnames(Buffers) = c("Site","Buffer1","Buffer2","Buffer3")

#Now add all these vars to site level vars data.frame

#Positive codes, richness and buffers can be added on the end

tmp = inner_join(SiteLevelVars,SiteCodes)
tmp = inner_join(tmp,Buffers)
tmp = inner_join(tmp,SiteRichness)

#remove all the site designations excep Any.Ac and num of designations.
tmp = tmp[-c(7:14, 16,17,19)]

#No add the sd/ranges from the plot level variables.
#First get the sd of pH, dbh etc

#missing plots

missing_sites = c( 3,15,17,38,46,46,46,59,59,  63,74,77, 83,90,98,98,98, 59,79,79,91,99,99,99,48,48,29,29,82,77)
missing_plots = c(14, 1,15, 9, 2, 8,16, 1, 8,  14,16, 1, 16, 2,11,12,14,  1, 2, 3,15,10,13,14, 7,10, 2, 1, 7,16)
missing = as.data.frame(cbind(missing_sites,missing_plots))
colnames(missing) = c("Site","Plot")

#remove missing plots
AllPlotsvars = anti_join(AllPlotsvars, missing, by = c("Site","Plot"))
#now if NA implies zero, not missing, so na.omit etc should be OK
#put zeros in dbh etc, because they are for no trees and we want to 
#count that in sd
AllPlotsvars$LiveBasalAreaYr2[is.na(AllPlotsvars$LiveBasalAreaYr2)]=0
AllPlotsvars$mean_dbh[is.na(AllPlotsvars$mean_dbh)]=0
AllPlotsvars$tree.density[is.na(AllPlotsvars$tree.density)]=0

hetero_vars = data.frame()

for (i in 1:103){
  site = AllPlotsvars%>%filter(Site == i)
  no_NVC = length(unique(site$ShortNVC))
  sd_pH = sd(site$pHYr2, na.rm = TRUE)
  sd_SOM = sd(site$SOMYr2,na.rm = TRUE)
  no_MSG = length(unique(site$MSG))
  sd_LBA = sd(site$LiveBasalAreaYr2,na.rm = TRUE)
  sd_meandbh= sd(site$mean_dbh,na.rm = TRUE)
  sd_treedensity= sd(site$tree.density,na.rm = TRUE)
  no_trees = sum(site$LiveBasalAreaYr2==0)
  vect = c(i,no_NVC, sd_pH, sd_SOM, no_MSG, sd_LBA, sd_meandbh, sd_treedensity,no_trees)
  hetero_vars = rbind(hetero_vars,vect)
 }

colnames(hetero_vars) = c("Site","no_NVC", "sd_pH", "sd_SOM", "no_MSG", "sd_LBA",
                          "sd_meandbh", "sd_treedensity","no_trees")


SiteLevelVars = inner_join(tmp,hetero_vars)
#Now completesitelevelvars has the site level variables 
#plus the sd of the plot level vars in each site, plus richness

# now add the sd of ellenbergs and mean of ellenbergs
#for each site the mean will use all values recorded including repeats
#because we want to include the effecs of the different assemblages across the wood.
#eg, if light loving plants in plot 1 and same ones in plot 5, that information
#must be retained.


ellen_vars = data.frame()

for (i in 1:130){
  site = veg_ellens%>%filter(SITE == i)
  sd_R = sd(site$R, na.rm = TRUE)
  mean_R = mean(site$R, na.rm = TRUE)
  sd_N = sd(site$N, na.rm = TRUE)
  mean_N = mean(site$R, na.rm = TRUE)
  sd_W = sd(site$W, na.rm = TRUE)
  mean_W =mean(site$W,na.rm = TRUE)
  sd_L = sd(site$L, na.rm = TRUE)
  mean_L = mean(site$L, na.rm = TRUE)
  vect = c(i,sd_R,mean_R,sd_N,mean_N,sd_W,mean_W,sd_L,mean_L)
  ellen_vars = rbind(ellen_vars,vect)
}
colnames(ellen_vars) = c("Site","sd_R","mean_R","sd_N","mean_N",
                         "sd_W","mean_W","sd_L","mean_L")
CompleteSiteLevelVars = inner_join(SiteLevelVars,ellen_vars)

#going to add mean dbh, mean pH to the complete set. 
#Then there are 3 sets of models, richnes vs
#1 -physical: alt, area/perim/E-N, no_designations, Any Anc, buffers, ave pH,ave dbh, ave live basal
#2 -hetero:number ANC, no +ve site codes, sd_ph, sd_meandbh, sd_LBA, 
#sd_tree density, no_trees(this is num of plots with no trees), num NVC
#3 -ellenergs - plant based traits that should reflect all of above.


means = data.frame()
for (i in 1:103){
  site = AllPlotsvars%>%filter(Site == i)
  meandbh = mean(site$mean_dbh)
  meanph = mean(site$pHYr2,na.rm = TRUE)
  meanSOM = mean(site$SOMYr2,na.rm = TRUE)
  meanLBA = mean(site$LiveBasalAreaYr2)
  tmp = c(i,meandbh,meanph,meanSOM,meanLBA)
  means = rbind(means,tmp)
}
  colnames(means) = c("Site","meandbh","meanph","meanSOM","meanLBA")

CompleteSiteLevelVars = inner_join(CompleteSiteLevelVars,means)

# add area/perim ratio
CompleteSiteLevelVars$area_ratio = CompleteSiteLevelVars$Area_ha*10000/CompleteSiteLevelVars$Perim_m

```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#Now first consider "physical" variables, select subset of CompleteSiteLevelVars

physical_vars = c("Site","Alt_m", "Area_ha","Perim_m" ,"area_ratio", "Easting" , "Northing","Buffer1","Buffer2" ,
                  "Buffer3", "Richness", "meandbh", "meanph" , "meanSOM" ,"meanLBA" )              
physical = CompleteSiteLevelVars%>%select(physical_vars)



#normalise to 1 so you can plot together
#remove site and richness
cut_physical = physical[-c(1,11)]
norm_physical = apply(cut_physical, 2, function(X) X/max(X))
norm_physical = as.data.frame(cbind(physical$Site, physical$Richness,norm_physical))
colnames(norm_physical) = c("Site","Richness","Alt_m","Area_ha","Perim_m","area_ratio",
                          "Easting", "Northing","Buffer1","Buffer2" ,
                           "Buffer3", "meandbh", "meanph" , "meanSOM" ,"meanLBA" ) 

# melt by hand, melt not doing the right thing
norm_physical_cut = norm_physical[,-c(1,2)]
R = norm_physical$Richness
Richness_rep = rep(norm_physical$Richness,13)
melted = melt(norm_physical_cut)
melted_physical = as.data.frame(cbind(Richness_rep,melted))

###select out subsets to plot otherwise cant see anything
set1 = melted_physical%>%filter(variable == "Alt_m" | variable == "Area_ha" |
                                  variable == "Perim_m" | variable == "area_ratio")

#pdf("../Results/set1.pdf")
ggplot(set1, aes(value,Richness_rep))+geom_point(aes(colour = variable))+
  geom_smooth(method = "lm", aes(colour = variable))
```

```{r}
set2 = melted_physical%>%filter(variable == "Easting" | variable == "Northing")

#pdf("../Results/set2.pdf")
ggplot(set2, aes(value,Richness_rep))+geom_point(aes(colour = variable))+
  geom_smooth(method = "lm", aes(colour = variable))
```

```{r}
set3 = melted_physical%>%filter(variable == "Buffer1" | variable == "Buffer2" | variable == "Buffer3")

#pdf("../Results/set3.pdf")
ggplot(set3, aes(value,Richness_rep))+geom_point(aes(colour = variable))+
  geom_smooth(method = "lm", aes(colour = variable))
```

```{r}
set4 = melted_physical%>%filter(variable == "meandbh" | variable == "meanph" | 
                                  variable == "meanSOM" | variable == "meanLBA")
#pdf("../Results/set4.pdf")
ggplot(set4, aes(value,Richness_rep))+geom_point(aes(colour = variable))+
  geom_smooth(method = "lm", aes(colour = variable))
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#now do heterogeneity variables,
#have dataframe called hetero_vars before joined to richness, but I'll subract from the complete
heterog_vars = c("Site","Richness","no_NVC", "Pos_Hetero_Index", "sd_pH", "sd_SOM", "no_MSG", "sd_LBA",
  "sd_meandbh", "sd_treedensity","no_trees")

heterog = CompleteSiteLevelVars%>%select(heterog_vars)
# replace NA by zero in Pos_Heter_Index column
heterog[is.na(heterog)] = 0
 #reshape

cut_heterog = heterog[-c(1,2)]
norm_heterog = apply(cut_heterog, 2, function(X) X/max(X))
norm_heterog = as.data.frame(cbind(heterog$Site, heterog$Richness,norm_heterog))
colnames(norm_heterog) = c("Site","Richness","no_NVC", "Pos_Hetero_Index", "sd_pH", "sd_SOM", "no_MSG", "sd_LBA",
                            "sd_meandbh", "sd_treedensity","no_trees") 

# melt by hand, melt not doing the right thing
norm_heterog_cut = norm_heterog[,-c(1,2)]
R = norm_heterog$Richness
Richness_rep = rep(norm_heterog$Richness,9)
melted = melt(norm_heterog_cut)
melted_heterog= as.data.frame(cbind(Richness_rep,melted))

set5 =  melted_heterog%>%filter(variable == "no_NVC" | variable == "no_MSG" | variable == "Pos_Hetero_Index" |
                                  variable == "sd_pH" | variable == "sd_SOM" )
#pdf("../Results/set5.pdf")
ggplot(set5, aes(value,Richness_rep))+geom_point(aes(colour = variable))+
  geom_smooth(method = "lm", aes(colour = variable))
```

```{r}
set6 = melted_heterog%>%filter(variable == "sd_meandbh" | variable == "sd_treedensity" |
                                 variable == "sd_LBA" | variable == "no_trees")
#pdf("../Results/set6.pdf")
ggplot(set6, aes(value,Richness_rep))+geom_point(aes(colour = variable))+
  geom_smooth(method = "lm", aes(colour = variable))
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
########## ELLENBERGS #############################

ellenberg_vars = c("Site","Richness","sd_R","mean_R","sd_N","mean_N",
                         "sd_W","mean_W","sd_L","mean_L")

ellenberg = CompleteSiteLevelVars%>%select(ellenberg_vars)
# replace NA by zero in Pos_Heter_Index column
#ellenberg[is.na(ellenberg)] = 0
#reshape

# melt by hand, melt not doing the right thing
ellenberg_cut = ellenberg[,-c(1,2)]
R = ellenberg$Richness
Richness_rep = rep(R,8)
melted = melt(ellenberg_cut)
melted_ellenberg= as.data.frame(cbind(Richness_rep,melted))


set7 =  melted_ellenberg%>%filter(variable == "sd_R" | variable == "sd_N" | 
                                  variable == "sd_W" | variable == "sd_L" )
#pdf("../Results/set7.pdf")
ggplot(set7, aes(value,Richness_rep))+geom_point(aes(colour = variable))+
  geom_smooth(method = "lm", aes(colour = variable))
```

```{r}
set8 = melted_ellenberg%>%filter(variable == "mean_R" | variable == "mean_N" |
                                 variable == "mean_W" | variable == "mean_L")
#pdf("../Results/set8.pdf")
ggplot(set8, aes(value,Richness_rep))+geom_point(aes(colour = variable))+
  geom_smooth(method = "lm", aes(colour = variable))

```

The plots show:
Positive correlation for the area of the wood. Larger woods have greater species richness in the 16 plots - possibly due to large species pool?
There is a negative correlation with altitude.

Positive correlation with northing - the northern woods are richer. Not sure why - are many of the southern woods beech??

Positive correlation with greater proportion of positive buffer - similar effect to having larger wood, species pool etc??

**Negative correlation with mean live basal area and mean dbh, as you would expect.**
**Negative correlation with soil organic matter - due to deep litter smothering** growth?? 

**Postive correlation with number of postive side codes recorded**
**Positive correlation with number of NVCs recorded.**

**Positive correlation with number of sites without trees**
Positive correlations with standard deviations of mean dbh, tree density and live basal area

**Positive correlation with increasing R and W ellenberg values**

This last correlation seems in agreement with the Boch paper and corresponds to the positive correlation of number of plots with no trees. As open areas are introduced, through forest management or the presence of tracks or riparian zones (all recorded in the Pos_Hetero_Index) light demanding species are seen, which add to the richness.

Possibly the positive correlation of the Ellenberg W corresponds to the correlation with the number of NVC codes or positive site codes. Both of these are reflecting different habitat types. 

The purpose of this exploration was a) to see which variables are having the strongest effect on the richness, and b)  to see what the effect o he different "groups" of variables - physical/heterogeneous/Ellenbergs.







