---
title: "ExaminingCovariatesVI"
author: "PetraGuy"
date: "27 March 2018"
header-includes: 
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	out.width="50%"
)
rm(list = ls())
cat("\014")
library(dplyr)
library(visreg) # for visreg plots 
library(ggplot2)
library(GGally) # for ggpairs
library(factoextra)
library(FactoMineR)# these two for PCA
library(corrplot)
library(car)
library(gridExtra)
library(reshape)
library(lattice)
```


```{r}
sitevars =  read.csv("../Data/CompleteSiteLevelVars.csv")
#toremove = c(23,53,74)
#sitevars = sitevars%>%filter(!Site %in% toremove)
colnames(sitevars)
```

```{r}
sitevars_reduced = subset(sitevars, select = -c(X,Site,No.of.Des, Any.Anc,Buffer1,Buffer2,sd_pH, 
                                                sd_SOM,sd_LBA, sd_meandbh,sd_treedensity,
                                                sd_R,mean_R, sd_N,mean_N , sd_W, mean_W,
                                                sd_L, mean_L ))
sitevars_all_lesRich = subset(sitevars_reduced, select = -c(Richness))
colnames(sitevars_reduced)
```
\clearpage
\newpage
\blandscape

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor) 
{
    usr <- par("usr"); on.exit(par(usr)) 
    par(usr = c(0, 1, 0, 1)) 
    r <- abs(cor(x, y)) 
    txt <- format(c(r, 0.123456789), digits=digits)[1] 
    txt <- paste(prefix, txt, sep="") 
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt) 
 
    test <- cor.test(x,y) 
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE, 
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " ")) 
 
    text(0.5, 0.5, txt, cex = cex ) 
    text(.8, .8, Signif, cex=cex, col=2) 
}

```

```{r out.width="592px", out.height="595px"}

pairs(sitevars_all_lesRich, lower.panel = panel.smooth, upper.panel = panel.cor)
      

       
```

Area,perimeter and area ratio, are correlated, as you’d expect. 
Also Buffer and Northing (0.52) Do Northern woods have larger buffers because there’s more room?

Also mean tree density, meanLBA, meanDBH no_trees, (0.306, -0.442, -0.417,) also as you’d expect, since all are also computationally similar. But meantreedensity is also correlated with  buffer, Northing and Easting (-0.38, -0.416, 0.336)

This implies that Eastern woods are denser while northern woods are more open, and woods with a larger buffer are less dense - yes, because they are northern.Buffer is correlated with northing (0.52), then what we are  seeing is the expression of the less dense northern woods.
\clearpage
\newpage
\elandscape

