---
title: "ExaminingCovariatesIV"
author: "PetraGuy"
date: "27 March 2018"
header-includes: 
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	out.width="50%"
)
rm(list = ls())
cat("\014")
library(dplyr)
library(visreg) # for visreg plots 
library(ggplot2)
library(GGally) # for ggpairs
library(factoextra)
library(FactoMineR)# these two for PCA
library(corrplot)
library(car)
```


```{r}
sitevars =  read.csv("../Data/SiteLevelVars.csv")
#toremove = c(23,53,74)
#sitevars = sitevars%>%filter(!Site %in% toremove)

```

```{r}
physical_df = sitevars%>%select("Alt_m", "Area_ha","Perim_m","Easting","Northing","Buffer3","meandbh","meanph","meanSOM","meanLBA","area_ratio")
hetero_df = sitevars%>%select("Pos_Hetero_Index","no_NVC","sd_pH", "sd_SOM" , "no_MSG", "sd_LBA", "sd_meandbh", "sd_treedensity","no_trees"  )
hetero_df[is.na(hetero_df)] = 0
physical_df = subset(physical_df, select = -c(Northing, Area_ha, Perim_m, meanLBA, Alt_m, Easting, meanph))
hetero_df = subset(hetero_df, select=-c(sd_meandbh,sd_treedensity,sd_SOM, sd_LBA, sd_pH, no_MSG, no_trees))
combo_df = as.data.frame(cbind(physical_df,hetero_df))
```

Combined data fram variables
```{r}
colnames(combo_df)
```

#Effect of each variable on richness using multiple linear regression, simple additive model.


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
combo_df_incRich = as.data.frame(cbind(sitevars$Richness, combo_df))
colnames(combo_df_incRich) = c("Richness", "Buffer3" , "meandbh" ,  "meanSOM"  , "area_ratio" , "Pos_Hetero_Index", "no_NVC" )
combo_model = lm(Richness ~., data = combo_df_incRich)

visreg(combo_model)
```


\clearpage
\newpage
\blandscape

#Colinearity of variables using pair plots

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="250%"}

ggpairs(combo_df, axisLabels="none",
        lower = list(continuous="smooth"),
        diag="blankDiag",
        upper = list(corSize=2,axisLabels='show'))+
theme(legend.position = "none",
      panel.grid.major = element_blank(),
      axis.ticks = element_blank(),
      panel.border = element_rect(linetype = "dashed", colour = "black", fill = NA))

```


\clearpage
\newpage
\elandscape


#PCA
##Physical variables

```{r}
combo_PCA = PCA(combo_df, scale.unit = TRUE, ncp = 11, graph = FALSE)
var = get_pca_var(combo_PCA)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
eig.val_physical = get_eigenvalue(combo_PCA)
eig.val_physical
```



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_eig(combo_PCA, addlabels = TRUE, ylim = c(0,30))
```
    
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_pca_var(combo_PCA, col.var = "cos2", repel = TRUE)
```

 
```{r echo=FALSE, message=FALSE, warning=FALSE}

corrplot(var$cos2)
```



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_cos2(combo_PCA, choice = "var")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_contrib(combo_PCA, choice = "var", axes = 1, top = 10)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
fviz_contrib(combo_PCA, choice = "var", axes = 2, top = 10)

```




fitall<- lm(formula = Richness ~  Buffer3 + meandbh + 
                meanSOM+area_ratio + Pos_Hetero_Index + no_NVC,
                data = combo_df_incRich)









